Mixing time estimation in reversible Markov chains from a single sample path

### 2015-06-01 intro-nips.tex
tweaks and comments in intro/abstract

In this paper we study the construction of fully empirical bounds on the mixing time of Markov chains based on a single sample path.
Let $(X_t)_{t=1,2,\dots}$ be an irreducible, aperiodic time-homogeneous Markov chain on a finite state space $[d]$ of $d$ states numbered by positive integers with transition matrix $\vP$. Under this assumption, the chain converges to its unique stationary \todoc{Maybe steady state is better?}
distribution $\vpi = (\pi_i)_{i=1}^d$ regardless of the distribution $\vq$ of $X_1$: 
$\lim_{t\to\infty} \Pr_{\vq}\Parens{X_t = i} = \lim_{t\to\infty} (\vq \vP^t)_i = \pi_i$, $1\le i \le d$.
The mixing time $\tmix$ of the Markov chain is the time that it takes for the chain to reach its stationary distribution, say,
\[
  \tmix
  :=
  \min\Braces{
    t \in \bbN :
    \sup_{\vq}
    \max_{A\subset [d]} |\textstyle\Pr_{\vq}\Parens{ X_t \in A } - \vpi(A)|
%    \Norm{
%      \vq \vP^t - \vpi
%    }_{\tv} 
    \leq 1/4
  }\,.
\]
Here $\vpi(A) = \sum_{i\in A} \pi_i$ is the total probability assigned to set $A$ by $\vpi$ and the supremum is over
all possible initial distributions $\vq$.
The problem studied in the paper is the construction of a confidence set $C_n = C_n(X_1,\dots,X_n,\delta) \subset [0,\infty]$ based on the knowledge of $X_1,\dots,X_n$ only that captures the mixing time $\tmix$ with a pre-specified probability $0\le \delta \le 1$. 
In the remainder of the introduction, we describe the background of why this problem might be interesting, followed by a 
brief description of our main results.

>>>>

In this paper we study the construction of fully empirical bounds on the mixing time of Markov chains based on a single sample path.
Let $(X_t)_{t=1,2,\dotsc}$ be an irreducible, aperiodic time-homogeneous Markov chain on a finite state space $[d]$ of $d$ states numbered by positive integers with transition matrix $\vP$. Under this assumption, the chain converges to its unique stationary \todoc{Maybe steady state is better?}
distribution $\vpi = (\pi_i)_{i=1}^d$ regardless of the distribution $\vq$ of $X_1$: 
$\lim_{t\to\infty} \Pr_{\vq}\Parens{X_t = i} = \lim_{t\to\infty} (\vq \vP^t)_i = \pi_i$, $1\le i \le d$.
The mixing time $\tmix$ of the Markov chain is the time that it takes for the chain to reach its stationary distribution, say,
\[
  \tmix
  :=
  \min\Braces{
    t \in \bbN :
    \sup_{\vq}
    \max_{A\subset [d]} |\textstyle\Pr_{\vq}\Parens{ X_t \in A } - \vpi(A)|
%    \Norm{
%      \vq \vP^t - \vpi
%    }_{\tv} 
    \leq 1/4
  }\,.
\]
Here $\vpi(A) = \sum_{i\in A} \pi_i$ is the total probability assigned
to set $A$ by $\vpi$, and the supremum is over
all possible initial distributions $\vq$.
The problem studied in the paper is the construction of a confidence
set $C_n = C_n(X_1,X_2,\dotsc,X_n,\delta) \subset [0,\infty]$ based on
the knowledge of $X_1,X_2,\dotsc,X_n$ only that captures the mixing
time $\tmix$ with a pre-specified probability $1-\delta \in (0,1)$. 
In the remainder of the introduction, we describe the background of why this problem might be interesting, followed by a 
brief description of our main results.
\todod{``might be interesting'' sounds weak.}
